---
title: "Extrapolation of chamber N$_2$O fluxes using Bayesian Kriging, and verification by eddy covariance: data processing and analysis"
author: "Peter E. Levy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, rendering, eval=FALSE}
library(rmarkdown)
library(here)
knitr::purl("vignettes/LincolnFluxDataAnalysis.Rmd",
  output = "vignettes/LincolnFluxDataAnalysis.R")
system.time(knitr::knit("LincolnFluxDataAnalysis.Rnw"))
tools::texi2pdf("LincolnFluxDataAnalysis.tex")
source("LincolnFluxDataAnalysis.R")
knitr::knit("LincolnFluxDataAnalysis.Rnw")
```




## Background
This document describes the processing of data for the paper \textit{Extrapolation of chamber N$_2$O fluxes using Bayesian Kriging, and verification by eddy covariance}, currently in preparation.  In the paper, we use a formal approach, Bayesian Kriging, for addressing the issue of upscaling chamber flux measurements to the larger domain measured by eddy covariance, so that the two methods can be compared on an equivalent basis.  The approach is related to Ordinary Kriging, in that it is a method for interpolating between data points, based on the observed spatial covariance structure, commonly shown graphically as a (semi-)variogram.  However, in the Bayesian form, rather than estimating the parameters of the variogram by least-squares regression, we put a represent them as a distribution, and update the distribution using the data and Bayes Theorem via Monte Carlo Markov Chain (MCMC) methods.  The advantage of this approach is that it makes spatial predictions, based on the observed chamber flux data, but accounts explicitly for the uncertainty in the spatial pattern across the domain. Where the area sampled is small compared to the domain of interest (typically \textless 1\% in the case of GHG measurements in agricultural fields), this is a major source of uncertainty which is generally ignored.

## How the code and documentation works
All the code is written in R with the ``literate programming'' package knitr.  The file LincolnFluxDataAnalysis.Rnw contains the R code and documentation in \LaTeX\, which is processed by the R command: \\
knitr::knit("LincolnFluxDataAnalysis.Rnw") \\
\\
to produce the numerical results, graphics and a .tex \LaTeX\ file.  The .tex file is then converted to (this) pdf document with the R command: \\
tools::texi2pdf("LincolnFluxDataAnalysis.tex") \\
\\
The R code is split into ``code chunks'', each of which can be set to not run with eval=FALSE to save time (particularly, readStackFootprint takes 30 mins to run, and interpolating over large or dense grids is slow).  Pre-calculated data is loaded from a .RData file at the start, and only re-calculated as needs be.  \\
\\
In brief, we first set the co-ordinates of the eddy covariance flux tower and the spatial prediction grid to interpolate chamber fluxes over.  We then read in all the various data sources, and do the necessary processing and merging.  The chamber fluxes are interpolated over the prediction grid using the krige.bayes function in the geoR package.  We extract the flux footprint over this prediction grid, and use this to get a footprint-weighted mean chamber flux.  This should produce a chamber-derived flux with the same spatial sampling characteristics as the eddy covariance measurement.  The uncertainty in this mean is given by the variability in the mean across all the simulations produced by krige.bayes.  The footprint-weighted chamber fluxes are plotted against the eddy covariance fluxes, and contrasted with the mean and uncertainties of the unweighted mean chamber flux (without interpolation).

## Packages required
The following R packages are used for graphics and spatial analysis.
```{r, clean up and load required packages}
rm(list=ls(all=TRUE))
here::i_am("vignettes/LincolnFluxDataAnalysis.Rmd")
# install.packages(c("Rmisc", "geoR", "rgdal"))
library(knitr)
library(ggplot2)
library(RColorBrewer)
library(ggmap)
library(sp)
library(rgdal)
library(raster)
library(openair)
library(geoR)
library(Rmisc)
Sys.setenv(TZ="UTC")
```

## Reading pre-calculated data or recalculating
If eval=FALSE for any of the later code chunks, we need to load pre-calculated data with the next code chunk.  If eval=TRUE for any of the later code chunks, the precalculated data in memory will be over-written (but not on disk with out doing the manual save, commented out in this code chunk).
```{r, load_precalculated_data}
#save(Fn2oCh, Fn2oChYork, Fn2oEC, Fco2EC, mergedEC_df, mergedChEC_df, stackFootprint, file="Fn2o_Brattleby.RData")
load("data/Fn2o_Brattleby.RData", verbose=TRUE)
```

## Defining tower and prediction grid
```{r, set_constants}
# OSGB 1936 / British National Grid
projOSGB <-  CRS("+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +datum=OSGB36 +units=m +no_defs")
# lat / lon
projlonlat <- CRS("+proj=longlat +datum=WGS84")

######### Set up prediction grid. #############
# Smaller resolution, higher computation time
res <- 10 # resolution for prediction grid, m
# Options for size and location of grid:
# 1. A specified extent - the CEH chambers bbox
extentCEHchambers <- extent(494247.6, 494291.6, 380905.1, 380945.1)
# 2. A specified extent - the CEH+York chambers bbox
extentAllchambers <- extent(494181.3, 494290.2, 380903.6, 381029.4)
# 3. A specified extent - the whole field bbox
extentField <- extent(494100, 494350, 380750, 381050)

# Or:
# 4. Square centred on flux tower, size given by
# grid_lth = length of side in metres
grid_lth   <- 100 # m
grid_hlth   <- grid_lth / 2
# 5. Offset from flux tower, given by shiftE & N
# Leave as zero if centred on tower
shiftE     <- -25 # m
shiftN     <- -25 # m

# define EC tower location
towerLon_degE <- -0.586180071
towerLat_degN <- 53.31718629
tower <- SpatialPoints(cbind(towerLon_degE, towerLat_degN))
proj4string(tower) <- projlonlat
towerOS <- spTransform(tower, projOSGB)
summary(towerOS)

extentTower <- extent(towerOS@coords[1,1]-grid_hlth+shiftE, towerOS@coords[1,1]+grid_hlth+shiftE,
                      towerOS@coords[1,2]-grid_hlth+shiftN, towerOS@coords[1,2]+grid_hlth+shiftN)

#### Here's where we choose option for grid size & location ####
# uncomment chosen option 1-5 (4/5 = w/o shifting)
#e <- extentCEHchambers
#e <- extentAllchambers
#e <- extentField
e <- extentTower
r <- raster(e, resolution = res, crs = projOSGB)

nSamplingSessions <- 12

# don't need this
r_wider_extent <- raster(nrows=60, ncols=50, xmn=494000, xmx=494500, ymn=380540, ymx=381200, crs=projOSGB)
r_field_extent <- raster(nrows=60, ncols=50, xmn=494100, xmx=494350, ymn=380750, ymx=381050, crs=projOSGB)
r_ECextent     <- raster(nrows=60, ncols=50, xmn=494240, xmx=494300, ymn=380900, ymx=380950, crs=projOSGB)

# function to add bounding box to plot
plot.bbox <- function(enew, lwd =1, col = 'black'){
  lines(c(enew@xmin, enew@xmin, enew@xmax, enew@xmax,
  enew@xmin),c(enew@ymin, enew@ymax, enew@ymax, enew@ymin,
  enew@ymin), lwd=lwd, col =col)
}
```

## Static chamber N$_2$O flux data, CEH Edinburgh
Static chamber flux measurments were carried out by CEH Edinburgh and Uni. York.  The CEH Edinburgh static chamber fluxes were calculated by the program RCflux (link to this here), and here we read in the output file from this program.  All files are stored in the GREENHOUSE project folder (P:\NEC04917_GREENHOUSE), currently run from \Sexpr{wd}.

As far as possible, we define functions to carry out separate tasks in the processing, to keep the program modular and clearer.  Here, we define a function to read and process the CEH Edinburgh chamber fluxes.  The function below reads a csv file, converts flux units to nmol m$^{-2}$\,s$^{-1}$, and merges fluxes with ancilliary data (containing spatial locations, measurement time, soil temperature etc.).  The function returns an object Fn2oCh (Flux of N2O from Chambers), which contains a dataframe df with one row per flux measurement.

```{r, read_CEH_chamber_data, eval=FALSE}
readFn2oCh <- function(){
  # read in CEH Edinburgh static chamber flux data from file
  df <- read.csv("../../RCflux/input/Brattleby/RCfluxOutput.csv")

  # convert umol to nmol m-2 s-1 in all flux columns
  colnames(df[,5:15])
  df[,5:15] <- df[,5:15]*1000

  # read in ancilliary data
  byInputFiles_df                <- read.csv("../../RCflux/input/Brattleby/ancilliaryData_byInputFiles.csv")
  byChamber_df                   <- read.csv("../../RCflux/input/Brattleby/ancilliaryData_byChamber.csv")
  bySamplingSession_byChamber_df <- read.csv("../../RCflux/input/Brattleby/ancilliaryData_bySamplingSession_byChamber.csv")
  names(df)

  df <- merge(df, byInputFiles_df, all.x = TRUE, sort = FALSE)
  df <- merge(df,     byChamber_df,    all.x = TRUE, sort = FALSE)
  df <- merge(df,     bySamplingSession_byChamber_df,
    by.x = c("date", "chamberID"),
    by.y = c("date", "chamberID"), all.x = TRUE, sort = TRUE)

  # convert date string to a real date/time
  df$TIMESTAMP <- paste(df$date, df$startTime_EChh)
  df$datech <- df$date
  df <- within(df, date_time <- as.POSIXct(strptime(TIMESTAMP, "%d/%m/%Y %H:%M")))
  df <- within(df, datect <- as.POSIXct(strptime(date, "%d/%m/%Y"), tz = "UTC"))
  df <- subset(df, gasName == "N2O") # remove CH4 data
  return(list(df=df))
}
Fn2oCh <- readFn2oCh()
```

```{r, plot_histogram}
ggplot(Fn2oCh$df, aes(x = flux_linear)) + geom_histogram(binwidth = 5) + facet_wrap( ~ datect, scales = "fixed")
```

flux_linear is the flux calculated by linear regression, as opposed to other regression methods, and should probably be replaced.  The histogram shows that there were 13 sampling days between 19/3/2014 and 14/4/2014.  Fluxes are roughly lognornal.

## Create SpatialPointsDataFrame
```{r, make_spdf_and_avg, eval=FALSE}
df2spdf <- function(df, projlonlat, projOSGB){
  # make into a SpatialPointsDataFrame
  spdf <- df
  coordinates(spdf) <- ~lon + lat
  proj4string(spdf) <- projlonlat
  spdf <- spTransform(spdf, projOSGB)
  return((spdf=spdf))
}
Fn2oCh$spdf <- df2spdf(Fn2oCh$df, projlonlat, projOSGB)
Fn2oCh$spdf$distFromTower <- spDistsN1(Fn2oCh$spdf, towerOS)
# remove chambers outwith the prediction grid (don't need to do this, but SkyLine is different)
Fn2oCh$spdf <- crop(Fn2oCh$spdf, r)

spatAvg <- function(spdf, r){
  # do spatial averaging
  rc <- rasterize(spdf, r, field = "flux_linear", fun = mean)
  plot(rc)
  # convert to a SpatialPointsDataFrame
  Fn2o_mean_spdf <- rasterToPoints(rc, spatial=TRUE)
  colnames(Fn2o_mean_spdf@data) <- "flux_linear"
  return((Fn2o_mean_spdf=Fn2o_mean_spdf))
}
Fn2oCh$Fn2o_mean_spdf <- spatAvg(Fn2oCh$spdf, r)
```





## Eddy covariance N$_2$O flux data, CEH Edinburgh
Some text.
```{r, readFn2oEC, eval=FALSE}
readFn2oEC <- function(){
  df <- read.csv("../../Data/Lincoln/EC/eddypro_Linc_essentials_2015-04-29T133448.csv",
     stringsAsFactors = FALSE, strip.white = TRUE, na.strings = c("NA","","-9999.0000000000000", "-9999") )
  length(df)
  for (i in 5:272){
    df[,i] <- suppressWarnings(as.numeric(df[,i]))
  }
  df$TIMESTAMP <- paste(df$date, df$time)
  df$datech <- df$date
  df <- within(df, date <- as.POSIXct(strptime(TIMESTAMP, "%Y-%m-%d %H:%M")))
  df <- within(df, datelt <- as.POSIXlt(strptime(TIMESTAMP, "%Y-%m-%d %H:%M")))
  df <- within(df, hour <- datelt$hour)

  df$ws <- df$wind_speed; df$wd <- df$wind_dir
  windRose(df)
  df$n2o_flux <- df$n2o_flux*1000 # convert umol to nmol m-2 s-1
  df$ru_n2o <- df$ru_n2o*1000 # convert umol to nmol m-2 s-1
  hist(df$n2o_flux)
  return(list(df=df))
}
Fn2oEC <- readFn2oEC()
```

## Eddy covariance CO$_2$ flux data, CEH Wallingford
Some text.
```{r, readFco2EC, eval=FALSE}
readFco2EC <- function(){
  df <- read.csv("../../Data/Lincoln/EC/eddypro_Brattleby_ConCrop_OilSee_essentials_2014-08-19T194333.csv",
     stringsAsFactors = FALSE, strip.white = TRUE, na.strings = c("NA","","-9999","-9999.0000000000000","-6999.0000000000000") )
  for (i in 5:274){
    df[,i] <- suppressWarnings(as.numeric(df[,i]))
  }
  df$TIMESTAMP <- paste(df$date, df$time)
  df$datech <- df$date
  df <- within(df, date <- as.POSIXct(strptime(TIMESTAMP, "%Y-%m-%d %H:%M")))
  df <- within(df, datelt <- as.POSIXlt(strptime(TIMESTAMP, "%Y-%m-%d %H:%M")))
  df <- within(df, hour <- datelt$hour)
  summary(df)

  df$ws <- df$wind_speed; df$wd <- df$wind_dir
  #windRose(df)
  #hist(df$wind_dir)
  return(list(df=df))
}
Fco2EC <- readFco2EC()
```

## Static chamber flux data, Uni. York
Some text.
```{r, readFn2oChYork, eval=FALSE}
readFn2oChYork <- function(){
  #N:\0Peter\prop\UKinverseFlux\GHG_TAP\skyline_vsEC\skyline.csv
  # read in Uni. York SkyLine flux data
  #load(file="C:/0Peter/prop/UKinverseFlux/GHG_TAP/skyline_vsEC/Fn2o_SkylineChambers.RData", verbose=TRUE)
  df <- read.csv("../../Data/Lincoln/skyline_vsEC/skyline.csv")
  df <- within(df, date   <- as.POSIXct(strptime(DATETIME, "%d/%m/%Y %H:%M")))
  df <- within(df, datelt <- as.POSIXlt(strptime(DATETIME, "%d/%m/%Y %H:%M")))
  df <- within(df, hour <- datelt$hour)
  df$chamberID <- df$CHANNEL
  df$chamberID <- df$chamberID + 100
  #df <- subset(df, N2O_nmolM2S > -10 & N2O_nmolM2S < 60)
  return(list(df=df))
  #save(df, file="C:/0Peter/prop/UKinverseFlux/GHG_TAP/skyline_vsEC/Fn2o_SkylineChambers.RData")
}
Fn2oChYork <- readFn2oChYork()
plot(N2O_nmolM2S ~ date, data=Fn2oChYork$df)
```

## Merge data frames
Some text.
```{r, merge data frames, eval=FALSE}
mergedEC_df <- merge(Fn2oEC$df,     Fco2EC$df,
  by.x = c("date"),
  by.y = c("date"), all.x = TRUE, sort = TRUE)

mergedChEC_df <- merge(Fn2oCh$df,     Fn2oEC$df,
  by.x = c("date_time"),
  by.y = c("date"), all.x = TRUE, sort = TRUE)

mergedChEC_df$ws <- mergedChEC_df$wind_speed; mergedChEC_df$wd <- mergedChEC_df$wind_dir
mergedChEC_df$chamberLoc <- "Field"
mergedChEC_df$chamberLoc[mergedChEC_df$chamberID > 16] <- "Skyline"
mergedChEC_df <- subset(mergedChEC_df, samplingSession.x <= nSamplingSessions)
```

## Eddy covariance footprint data, CEH Edinburgh
Some text.
```{r, readStackFootprint, eval=FALSE}
# takes 30 mins to run
readStackFootprint <- function(nSamplingSessions){
  # read in the footprint raster stack
  load("../../Data/Lincoln/EC/footprintStack_Brattleby_index.RData", verbose=TRUE)
  s <- stack("../../Data/Lincoln/EC/footprintStack_Brattleby.tif")
  # initialise stack for sampling session means
  sfp <- s[[1:nSamplingSessions]]
  sfp <- setValues(sfp, NA)
  sfp

  # find the indices in the footprint stack for each chamber measurement
  ind <- match(Fn2oCh$df$date_time, index_df$datect, nomatch = NA_integer_, incomparables = NULL)
  Fn2oCh$df$samplingSession.y
  names(Fn2oCh$df)

  # get the indices in the footprint stack for all chamber measurements in each sampling session
  # and average the footprint over these times
  # could be weighted towards the more common times without getting just the unique list
  for (iSamplingSession in 1:nSamplingSessions){
    #iSamplingSession <- 1
    sessionIndices <- ind[Fn2oCh$df$samplingSession.y == iSamplingSession]
    sessionIndices <- subset(sessionIndices, !is.na(sessionIndices))
    sessionIndices <- unique(sessionIndices)
    sfp[[iSamplingSession]] <- mean(s[[sessionIndices]])
  }
  plot(sfp[[1:nSamplingSessions]])
  cellStats(sfp[[1]], sum)
  return(list(sfp=sfp))
}
stackFootprint <- readStackFootprint(nSamplingSessions)
```

## Interpolate chamber fluxes with Bayesian Kriging
Some text.
```{r, interpolateChamberFlux}
pred_grid <- rasterToPoints(r, spatial=TRUE)
interpolateChamberFlux <- function(spdf, r, pred_grid = rasterToPoints(r, spatial=TRUE), nSamplingSessions){
  #spdf <- Fn2oCh$spdf
  grid_coords <- rasterToPoints(r)
  s_Fn2o_interp <- stack()
  # initialise vector for sd
  Fn2o_ch_interp_sd <- seq(1, nSamplingSessions)

  par(mfrow = c(4, 2))
  for (isamplingSession in 1:nSamplingSessions){  # max(spdf$samplingSession.x)
    #isamplingSession <- 1
    sdf <- subset(spdf, samplingSession.x == isamplingSession)
    geo_Fn2o <- as.geodata(sdf, data.col = 7, covar.col=40)
    maxdist  <- variog(geo_Fn2o)$max.dist
    seqphi_sparse <- seq(0,2*maxdist,l=)
    # Now we do the analysis:
    geo_Fn2o_bayes <- krige.bayes(
                          geo_Fn2o,
                          loc=grid_coords,
                          model=model.control( trend.d="cte", trend.l="cte", cov.model="matern",
                          # model=model.control( trend.d=trend.spatial(~ndvi, geo_Fn2o),
                          #                     trend.l=trend.spatial(~ndvi, pred_grid),
                          #                     cov.model="matern",
                                               kappa=0.5, aniso.pars=NULL, lambda=1 ),
                          prior=prior.control( beta.prior="flat",
                                               sigmasq.prior="reciprocal",
                                               phi.prior="uniform", phi.discrete=seqphi_sparse,
                                               tausq.rel.prior="fixed", tausq.rel=0 ),
                          output=output.control( n.posterior=100, messages=FALSE )
                                   )
    colMeans(geo_Fn2o_bayes$posterior$sample)
    rp <- setValues(r, geo_Fn2o_bayes$predictive$mean)
    rp <- flip(rp, direction = "y")
    #plot(rp, col=terrain.colors(21))
    #s_Fn2o_interp[[isamplingSession]] <- rp
    s_Fn2o_interp <- stack(s_Fn2o_interp, rp)
    Fn2o_ch_interp_sd[isamplingSession] <- sd(geo_Fn2o_bayes$predictive$sim.means)
    # We create the maps
    image  (geo_Fn2o_bayes,
            val=geo_Fn2o_bayes$predictive$mean,
            xlim=c(rp@extent@xmin,rp@extent@xmax), ylim=c(rp@extent@ymin-10,rp@extent@ymax),
            col=terrain.colors(21), xlab="lon", ylab="lat")
    points (geo_Fn2o, cex.max = 1, col = "black", add=T, pt.divide="equal" )
    contour(geo_Fn2o_bayes, add=T, nlev=11)
    legend.krige(val=geo_Fn2o_bayes$predictive$mean,
                 col=terrain.colors(21),
                 x.leg=c(rp@extent@xmin,rp@extent@xmax), y.leg=c(rp@extent@ymin-10,rp@extent@ymin-5) )
  }

  par(mfrow = c(4, 4))
  for (i in (1:16)) {
    image  (geo_Fn2o_bayes,
            val=geo_Fn2o_bayes$predictive$simulations[,i],
            xlim=c(rp@extent@xmin,rp@extent@xmax), ylim=c(rp@extent@ymin-10,rp@extent@ymax),
            col=terrain.colors(21), xlab="lon", ylab="lat")
    points (geo_Fn2o, cex.max = 0.5, col = "black", add=T, pt.divide="equal" )
    legend.krige(val=sqrt(geo_Fn2o_bayes$predictive$variance),
                 col=terrain.colors(21),
                 x.leg=c(rp@extent@xmin,rp@extent@xmax), y.leg=c(rp@extent@ymin-10,rp@extent@ymin-5) )
  }
  names(s_Fn2o_interp)[] <- "F_n2o_pred"
  return(list(s_Fn2o_interp=s_Fn2o_interp, Fn2o_ch_interp_sd=Fn2o_ch_interp_sd))
}
i <- interpolateChamberFlux(Fn2oCh$spdf, r, pred_grid, nSamplingSessions)
s_Fn2o_interp <- i$s_Fn2o_interp
Fn2o_ch_interp_sd <- i$Fn2o_ch_interp_sd
plot(s_Fn2o_interp)
```
